Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-06-06 analysis_parameters.tex
first general revision.
finished attenuation relationship section

\section{Analysis Parameters}
\subsection{Attenuation Relationship}

The choice of the ground motion attenuation relationship, as a highly influential factor, is of a great importance in seismic hazard analysis. \citet{Zafarani2014} used 163 free-field acceleration time histories recorded at epicentral distance of up to 200 km from 32 earthquakes to investigate the predictive capabilities of the local, regional, and next generation attenuation (NGA) ground-motion prediction equations and determined their applicability for northern Iran. After evaluating different Ground Motion Prediction Equations, \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008} represented suitable performance for  PGA  with LLH (Log-Likelihood method) score of 1.54, 1.55, and 1.59. Mean of the mentioned attenuation relationships (not shown here), are very close together, especially at higher magnitude. Using \citet{Scherbaum2009} approach and \citet{Zafarani2014} coefficients we calculated the logic tree weights as 0.3376, 0.3354, and 0.3270, respectively. The epistemic uncertainty in attenuation relationship is being considered through using the logic tree approach. Having close mean value and similar logic tree weights, logic tree approach will have a minor effect in considering the epistemic uncertainty. In this study instead of using  different GMPEs we use the best model of attenuation relationship for PGA \citep{Zafarani2014} as well $\pm$ standard deviation. \\

\noindent
The attenuation relationship is:

\begin{equation}
\begin{split}
ln\ (Y) = b_1 + b_2(M_w-6) + b_3( M_w-6)^{2}+  \\ 
b_5ln\ r + b_V \ ln(V_S/V_A) \  with \  r= \sqrt{R{r^2_{cl} + h^2}}  
\end{split}
\end{equation}

@@

\subsection{Catalog Completness}
For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor.  \citet{Frankel1995} plotted the cumulative number of events against time for different regions. He assumes that from the point that the line become linear, the catalog is complete. This approach is similar to the MAXC method of ZMAP software \citep{Wiemer2001}, where the completeness treshold is the maximum value of the first derivative of the frequency-magnitude curve. Using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and also frequency magnitude distribution approach in software ZMAP, \citet{Zare2014} reported the catalog completeness for the study regions. World widely large earthquakes were routinely located after increasing number of seismic stations establishment in the early 1900s \citep{Shearer2009}. Up to 1961 these data formed the early instrumental period. In 1961 the Worldwide Standardized Seismograph Network (WWSSN) was stablished. The record of these seismographs considerably improved the seismic catalogs in different part of the world \citep{Shearer2009}. Fig.~\ref{fig:completness_scatter} shows the magnitude distribution of events with respect to the time of occurrence of the events.

@@

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.28]{figures/pdf/comp_test_all_mag.pdf} 
\caption{Magnitude-time distribution of earthquakes in the study regions. a-b) Alborz, c-d) Azerbaijan, e-f)Kopek Dagh, g-h)Kopeh Dagh, i-j)Zagros}
\label{fig:comp_test_all_mag}
\end{figure*}

>>>>

\section{Analysis Parameters}

% \subsection{Attenuation Relationship}

% The choice of the ground motion attenuation relationship, as a highly influential factor, is of a great importance in seismic hazard analysis. \citet{Zafarani2014} used 163 free-field acceleration time histories recorded at epicentral distance of up to 200 km from 32 earthquakes to investigate the predictive capabilities of the local, regional, and next generation attenuation (NGA) ground-motion prediction equations and determined their applicability for northern Iran. After evaluating different Ground Motion Prediction Equations, \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008} represented suitable performance for  PGA  with LLH (Log-Likelihood method) score of 1.54, 1.55, and 1.59. Mean of the mentioned attenuation relationships (not shown here), are very close together, especially at higher magnitude. Using \citet{Scherbaum2009} approach and \citet{Zafarani2014} coefficients we calculated the logic tree weights as 0.3376, 0.3354, and 0.3270, respectively. The epistemic uncertainty in attenuation relationship is being considered through using the logic tree approach. Having close mean value and similar logic tree weights, logic tree approach will have a minor effect in considering the epistemic uncertainty. In this study instead of using  different GMPEs we use the best model of attenuation relationship for PGA \citep{Zafarani2014} as well $\pm$ standard deviation. \\

% \noindent
% The attenuation relationship is:

% \begin{equation}
% \begin{split}
% ln\ (Y) = b_1 + b_2(M_w-6) + b_3( M_w-6)^{2}+  \\ 
% b_5ln\ r + b_V \ ln(V_S/V_A) \  with \  r= \sqrt{R{r^2_{cl} + h^2}}  
% \end{split}
% \end{equation}

@@

\subsection{Catalog Completness}

For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor.  \citet{Frankel1995} plotted the cumulative number of events against time for different regions. He assumes that from the point that the line become linear, the catalog is complete. This approach is similar to the MAXC method of ZMAP software \citep{Wiemer2001}, where the completeness treshold is the maximum value of the first derivative of the frequency-magnitude curve. Using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and also frequency magnitude distribution approach in software ZMAP, \citet{Zare2014} reported the catalog completeness for the study regions. World widely large earthquakes were routinely located after increasing number of seismic stations establishment in the early 1900s \citep{Shearer2009}. Up to 1961 these data formed the early instrumental period. In 1961 the Worldwide Standardized Seismograph Network (WWSSN) was stablished. The record of these seismographs considerably improved the seismic catalogs in different part of the world \citep{Shearer2009}. Fig.~\ref{fig:completness_scatter} shows the magnitude distribution of events with respect to the time of occurrence of the events.

@@

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.28]{figures/pdf/figure-07.pdf} 
\caption{Magnitude-time distribution of earthquakes in the study regions. a-b) Alborz, c-d) Azerbaijan, e-f)Central, g-h)Kopeh Dagh, i-j)Zagros}
\label{fig:comp_test_all_mag}
\end{figure*}

### 2016-06-06 composition.tex
finished attenuation relationship section

Selection and declustering was also done in consistence with the original catalog, using the method of \citet{Gardner1974}. In summary, we obtained the whole catalog version of \citet{Zare2014}, removed the events prior to 2000, complemented with all $M>3$ events from 2000 to 2015 from IIEES, and finally declustered the new complemented catalog removing foreshocks and aftershocks. The resulting set of events compiled for this study is shown in Fig.~\ref{fig:catalog}.

>>>>

Selection and declustering was also done in consistence with the original catalog, using the method of \citet{Gardner1974}. In summary, we obtained the whole catalog version of \citet{Zare2014}, removed the events from 2000 and forward, complemented the catalog with all $M>3$ events up to 2015 from IIEES, and finally declustered the new complemented catalog removing foreshocks and aftershocks. The resulting set of events compiled for this study is shown in Fig.~\ref{fig:catalog}.

### 2016-06-06 introduction.tex
finished attenuation relationship section

Most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to data availability. We use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Boyd2008, Kalkan2009, Moschetti2014}, including Iran \citep{Khodaverdian_2016_BSSA}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Fig.~\ref{fig:iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region \citep{Zafarani2014, Kalkan2004}, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions, considering different magnitude thresholds. We compare our results to equivalent estimations (hazard curves) available in the literature for selected locations and conclude with a discussion in light of other previous hazard studies.

>>>>

Most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to data availability. We use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Cao1996, Akinci2004, Kalkan2009}, including Iran \citep{Khodaverdian_2016_BSSA}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Fig.~\ref{fig:iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions, considering different magnitude thresholds. We compare our results to equivalent estimations (hazard curves) available in the literature for selected locations and conclude with a discussion in light of other previous hazard studies.
